---
title: "Exercise 3"
author: "Arthur Junges Schmidt"
date: "`r as.character(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  documentclass: report
  html_notebook: default
  html_document:
    df_print: paged
geometry: margin=2cm
---
(a) How many regressor variables are in this model?
$$\text{Number of Variables = 3}$$

(b) If the error sum of squares is 307 and there are 15 observations, what is the estimate of $\sigma^{2}$?

$$\sigma^{2} = \frac{SSe}{N-1}$$

```{r}
SSe<- 307
N <- 15

Variance <- SSe/(N - 1) 
print(Variance)
```

(c) What is the standard error of the regression coefficient $\hat\beta_{1}$?

Variance $\sigma^{2}$ was calculated on question B, with value `r round(Variance, 3)`. The equation to discover the error of the regression coefficient is $$SE(\hat\beta_{1}) = \sqrt{\frac{\hat{\sigma}^{2}}{\sum_{i=1}^n  ( x_{i}- \overline{x})^{2} }}$$

Since the value on position [3,3] on the matrix is represents $$\sum_{i=1}^n  ( x_{i}- \overline{x})^{2}$$,
then the $SE(\hat\beta_{1})$ is $\sigma^{2}\times 0.0009108 =$ `r Variance * 0.0009108`

```{r}




```

