---
title: "Exercise 6"
author: "Arthur Junges Schmidt"
date: "`r as.character(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  documentclass: report
  html_notebook: default
  html_document:
    df_print: paged
geometry: margin=2cm
  
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
Original_Data <- read_csv("Problemset 2 data for problem 6 DaimlerChrysler.csv", 
    skip = 1)

Original_Data <- Original_Data[-(22:34),] ## Taking out unnecessary rows 
```

The data which can be found in a separate spreadsheet provides the highway gasoline mileage
test results for 2005 model year vehicles from DaimlerChrysler.
(1) Fit a multiple linear regression model to these data to estimate gasoline mileage that uses the
following regressors: cid, rhp, etw, cmp, axle, n/v

```{r}

Linear_Model <- lm(formula = mpg ~ cid + rhp + etw + cmp + axle + 
                     `n/v`, 
                   data = Original_Data);
summary(Linear_Model)

```



(2) Estimate $\sigma^{2}$ and the standard errors of the regression coefficients.

$\sigma^{2}$ values:
```{r echo=FALSE}
diag(vcov(Linear_Model))
```

Standard errors:
```{r echo=FALSE}
summary(Linear_Model)$coefficients[,2]
```

(3) Test for significance of regression using $\alpha$ = 0.05. What conclusions can you draw?
Only the intercept and the 'etw' and 'axle' variables have a significance of 95% or higher. This leads to the conclusion that, with 95% certainty, only the axle ration and the equivalent test weight variables are significant to explain the mileage of the vehicles on the data set.

(4) Find the t-test statistic for each regressor. Using $\alpha$ = 0.05, what conclusions can you draw? Does each regressor contribute to the model?

T-test statistic for each regressor:
```{r echo=FALSE}
summary(Linear_Model)$coefficients[,3]
```

As on the question before, the only 2 variables that have a significant t-test are 'etw' and 'axle'. Not all the regressor contributes to the model since some of them add more error into it.


(5) Find 99% confidence intervals on the regression coefficients.
```{r echo=FALSE}
confint.lm(Linear_Model, level = 0.99)
```

(6) Plot residuals versus $\hat{Y}$ and versus each regressor. Discuss these residual plots.
```{r fig.height=15, fig.width=10}
par(mfrow = c(5,2))
plot(x = Linear_Model$model$cid, y = resid(Linear_Model), xlab = "cid", ylab = "Residuals")
plot(x = Linear_Model$model$rhp, y = resid(Linear_Model), xlab = "rhp", ylab = "Residuals")
plot(x = Linear_Model$model$etw, y = resid(Linear_Model), xlab = "etw", ylab = "Residuals")
plot(x = Linear_Model$model$cmp, y = resid(Linear_Model), xlab = "cmp", ylab = "Residuals")
plot(x = Linear_Model$model$axle, y = resid(Linear_Model), xlab = "axle", ylab = "Residuals")
plot(x = Linear_Model$model$`n/v`, y = resid(Linear_Model), xlab = "n/v", ylab = "Residuals")
plot(Linear_Model)

```

